---
title: 'HW 5: Correlations and linear models. Tests for categorial variables'
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(vcd) # for mosaic plots
```

## 1. Vowel reduction in Russian
Pavel Duryagin ran an experiment on perception of vowel reduction in Russian language. The dataset `shva` includes the following variables:  
_time1_ - reaction time 1  
_duration_ - duration of the vowel in the stimuly (in milliseconds, ms)  
_time2_ - reaction time 2  
_f1_, _f2_, _f3_ - the 1st, 2nd and 3rd formant of the vowel measured in Hz (for a short introduction into formants, see [here](https://home.cc.umanitoba.ca/~krussll/phonetics/acoustic/formants.html))  
_vowel_ - vowel classified according the 3-fold classification (_A_ - _a_ under stress, _a_ - _a/o_ as in the first syllable before the stressed one, _y_ (stands for shva) - _a/o_ as in the second etc. syllable before the stressed one or after the stressed syllable, cf. _g_[_y_]_g_[_a_]_t_[_A_]_l_[_y_] _gogotala_ `guffawed').  
In this part, we will ask you to analyse correlation between f1, f2, and duration.
The dataset is available [https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt](here).
### 1.0 
Read the data from file to the variable `shva`.
```{r 1.0}
shva <- read_tsv('https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/duryagin_ReductionRussian.txt')
shva
```

### 1.1 
Scatterplot `f1` and `f2` using `ggplot()`. 
Design it to look like the [following](https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/img/duryagin1.png).
```{r 1.1}
col <- c("blue", "red", "green")[factor(shva$vowel)]
shva %>% ggplot(aes(f2, f1, colour = vowel)) + geom_point() + scale_x_reverse() + scale_y_reverse()
```

### 1.2 
Plot the boxplots of `f1` and `f2` for each vowel using `ggplot()`.
Design it to look like [this](https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/img/duryagin2.png) and [this](https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/img/duryagin3.png).
```{r 1.2}
# f1 boxplot
shva %>% ggplot(aes(vowel, f1)) + coord_flip() + geom_boxplot(aes(fill=vowel))
# f2 boxplot
shva %>% ggplot(aes(vowel, f2)) + coord_flip() + geom_boxplot(aes(fill=vowel))
```

### 1.3 
Calculate Pearson's correlation of `f1` and `f2` (all data)
```{r 1.3}
cor.test(shva$f1, shva$f2)
```

### 
1.4 Calculate Pearson's correlation of `f1` and `f2` for each vowel
```{r 1.4}
A_var <- filter(shva, vowel == "A")
a_var <- filter(shva, vowel == "a")
y_var <- filter(shva, vowel == "y")
cor.test(A_var$f1, A_var$f2)
cor.test(a_var$f1, a_var$f2)
cor.test(y_var$f1, y_var$f2)
```

## 2 Linear regressions

### 2.1.1 
Use the linear regression model to predict `f2` by `f1`.
```{r 2.1.1}
f2_pred <- lm(data = shva, f2 ~ f1)
summary(f2_pred)
```

### 2.1.2 
Write down the equation for f2 using coefficients from the model (e.g. $y =  b + kx$)
```{r 2.1.2}
print("f2 = 1639.70215 - 0.42875 * f1")
```

### 2.1.3 
Provide the adjusted R$^2$
```{r 2.1.3}
summary(f2_pred)$adj.r.squared
```

### 2.1.4 
Add the regression line in the scatterplot 1.1.
```{r 2.1.4}
col <- c("blue", "red", "green")[factor(shva$vowel)]
shva %>% ggplot(aes(f2, f1, colour = vowel)) + geom_point() + scale_x_reverse() + scale_y_reverse() + geom_smooth(method='lm', formula= y~x, inherit.aes = FALSE, aes(f2, f1))
```

### 2.1.5 
Make a scatter plot for `f1` and `f2` grouped by vowels. 
Use `ggplot()` and `facet_wrap()`.

```{r 2.1.5}
col <- c("blue", "red", "green")[factor(shva$vowel)]
shva %>% ggplot(aes(f2, f1, colour = vowel)) + geom_point() + facet_wrap(vars(shva$vowel))
```

### 2.2.1 
Use the linear regression model to predict `f2` by `f1` and `vowel`.
```{r 2.2.1}
f2_pred_vow <- lm(data = shva, f2 ~ f1 + vowel)
summary(f2_pred_vow)
```

### 2.2.2 
What is the intercept of the model?
```{r 2.2.2}
summary(f2_pred_vow)$coefficients[1, 1]
```

### 2.2.3 
Provide the adjusted R$^2$
```{r 2.2.3}
summary(f2_pred_vow)$adj.r.squared
```

### 2.2.4 
Write down your general conclusions about the relationship between `f1`, `f2`, and `vowels`.

```
Коэффициент детерминации для модели с константой принимает значения от 0 до 1. Чем ближе значение коэффициента к 1, тем сильнее зависимость. Для приемлемых моделей предполагается, что коэффициент детерминации должен быть хотя бы не меньше 50%. Модели с коэффициентом детерминации выше 80 % можно признать достаточно хорошими. Значение коэффициента детерминации 1 означает функциональную зависимость между переменными.
Здесь же мы наблюдаем, что коэффициент детерминации меньше 50%, что означает, что f2 зависит от f1 и vowel незначительно.
```

## 3. Dutch causative constructions

When the Dutch use two near-synonymous periphrastic causative verbs, *doen* and *laten*?
```
       De politie deed/liet de auto stoppen.
  lit. the police did/let the car stop 
       'The police stopped the car'
```

This is a data set on two rival constructions with *doen* and *laten* sampled from the newspaper corpora. The data frame includes 500 observations on the following 7 variables:

* `Aux` -- verb: a factor with levels `doen` and `laten`  
* `CrSem` -- the semantic class of the Causer: a factor with levels `Anim` (animate) and `Inanim` (inanimate)  
* `CeSem` -- the semantic class of the Causee: a factor with levels `Anim` (animate) and `Inanim` (inanimate)  
* `CdEvSem` -- the semantic domain of the caused event expressed by the Effected Predicate: a factor with levels `Ment` (mental) and `NonMent` (e.g. physical or social)  
* `CeSynt` -- the syntactic status of the Causee: a factor with levels `Clause`, `Impl` (implicit, not expressed), `NP` (noun phrase), `PP` (prepositional phrase)    
* `EPTrans` -- transitivity or intransitivity of the effected predicate, a factor with two levels `Tr` and `Intr`  
* `Country` -- a factor with levels `BE` (Belgium) and `NL` (Netherlands)  
* `Domain` -- a factor with four levels for newspaper domains.    

Data from Natalya Levshina's `RLing` package available (here)[https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv]
Read more on the constructions in [Levhina, Geerarts, Speelman 2014](https://www.academia.edu/7287585/Dutch_causative_constructions_Quantification_of_meaning_and_meaning_of_quantification_with_Dirk_Geeraerts_and_Dirk_Speelman_).

### 3.0 
Read the data from file to the variable `d_caus`.
```{r 3.0}
d_caus <- read.csv("https://raw.githubusercontent.com/agricolamz/2018-MAG_R_course/master/data/dutch_causatives.csv")
summary(d_caus)
```

### 3.1 
We are going to test whether the association between `Aux` and other categorical variables (`Aux` ~ `CrSem`, `Aux` ~ `CeSem`, etc) is statistically significant. The assiciation with which variable should be analysed using Fisher's Exact Test and not using Pearson's Chi-squared Test? Is this association statistically significant?

```{r 3.1}
chisq.test(d_caus$Aux, d_caus$CrSem)
chisq.test(d_caus$Aux, d_caus$CeSem)
chisq.test(d_caus$Aux, d_caus$CdEvSem)
chisq.test(d_caus$Aux, d_caus$CeSynt)
chisq.test(d_caus$Aux, d_caus$EPTrans)
chisq.test(d_caus$Aux, d_caus$Country)
chisq.test(d_caus$Aux, d_caus$Domain)
#Тест Фишера нам потребуется, чтобы проанализировать связь с CeSynt
fisher.test(d_caus$Aux, d_caus$CeSynt)
```
P-value в этом случае будет крайне небольшим, что означает, что эта зависимость является значимой.


### 3.2. 
Test the hypothesis that `Aux` and `EPTrans` are not independent with the help of Pearson's Chi-squared Test. 
```{r 3.2}
Aux_assoc <- chisq.test(d_caus$Aux, d_caus$EPTrans)
Aux_assoc 
```

Aux и EPTrans не являются независимыми величинами, так как p-value меньше пяти процентов.

### 3.3 
Provide expected frequencies for Pearson's Chi-squared Test of `Aux` and `EPTrans` variables.
```{r 3.3}
Aux_assoc$expected
```

### 3.4. 
Calculate the odds ratio for observed frequencies of `Aux` and `EPTrans`
For 2×2 contigency table
$$\begin{matrix}
a & b \\
c & d
\end{matrix}
$$
one can find *odds ratio* as $(a/c)/(b/d)$.
```{r 3.4}
a = Aux_assoc$observed[1, 1]
b = Aux_assoc$observed[1, 2]
c = Aux_assoc$observed[2, 1]
d = Aux_assoc$observed[2, 2]
odds_ratio = (a/c) / (b/d)
odds_ratio
```

### 3.4.1 
Find odds ratio for expected frequencies of `Aux` and `EPTrans`

```{r 3.4.1}
a = Aux_assoc$expected[1, 1]
b = Aux_assoc$expected[1, 2]
c = Aux_assoc$expected[2, 1]
d = Aux_assoc$expected[2, 2]
expected_or = (a/c)/(b/d)
expected_or
```
What can you say about odds ratio of expected frequencies for arbitrary data?

Отношение шансов равное 1 означает, что исследуемое событие обладает равными шансами в обеих группах. Отношение шансов превышающее 1 означает, что событие имеет больше шансов произойти в первой группе. И отношение шансов не превышающее 1 свидетельствует о том, что событие имеет меньше шансов в первой группе.
Наше ожидаемое отношение шансов равно единице, что означает их равную вероятность.

### 3.5 
Calculate effect size for this test using Cramer's V (phi).
```{r 3.5}
install.packages('lsr')
library(lsr)
cramersV(Aux_assoc$observed)
```

### 3.6. 
Report the results of independence test using the following template:
```
V Крамера = 0.1691587, это небольшой результат, что говорит о незначительном эффекте.
```

### 3.7 
Visualize the distribution using mosaic plot.
Use `mosaic()` function from `vcd` library.
```{r 3.7}
library(vcd)
mosaic(~ EPTrans + Aux, data=d_caus, shade=TRUE, legend=TRUE)
```

Below is an example of how to use mosaic() with three variables.
```{r 3.7.1}
mosaic(~ Aux + CrSem + Country, data=d_caus, shade=TRUE, legend=TRUE)
```

### 3.8 
Why is it not recommended to run multiple Chisq tests of independence on different variables within your dataset whithout adjusting for the multiplicity? (i.e. just testing all the pairs of variables one by one)  
```
Это связано с тем, что для каждого отдельного теста есть своя вероятность ошибки, и чем большее количество тестов мы проведём, тем выше будет вероятность ошибки в целом, а, значит, тем менее точными будут данные. 
```
Using multiple logistic regression and classiﬁcation trees, the study explores the conceptual diﬀerences between the constructions
### 3.9 
Provide a short text (300 words) describing the hypothesis of this study and the results of your analysis.
```{r 3.9}
print('В этом исследовании мы пытаемся разобраться, в каком случаев голландском языке употребляется глагол "doen", а в каком случае — глагол "laten" (примерноаналогичные английским "do" и "let"). Для исследования мы используем датафрейм, составленные при помощи газетного корпуса. В этом датафрейме у нас представлены 500 наблюдений с восемью переменными: Aux — это собственно глаголы "doen" и "laten"; CrSem — это семантический класс того, кто совершает действие (одушевлённый или неодушевлённый); CeSem — это семантический класс того, над чем совершают действие (одушевлённый или неодушевлённый); CdEvSem — семантическая сфера события (ментальная или нементальная); CeSynt — синтаксический класс того, над чем совершают действие (клауза, имплицитный, NP и PP); EPTrans — переходность или непереходность глагола; Country — Бельгия или Нидерланды; Domain — четыре различных газетных домена. Мы исследовали с помощью Хи-квадрат теста связь между этими переменными и переменной Aux. Для CeSynt пришлось использовать точный тест Фишера. Точный тест Фишера показал, что синтаксичееский класс может влиять на использование этих глаголов. С помощью хи-квадрата и отношения шансов нам также удалось выяснить, что "doen" чаще, чем "laten", используется с непереходными глаголами. Мы также подсчитали V Крамера, которая показала, что существенной связи между использованием того или иного глагола и переходностью/непереходностью нет. Благодаря всем использованным в работе тестам нам удалось выяснить много важной статистической информации. Такие тесты собранной информации были бы полезны во многих лингвистических исследованиях, так как это позволит более точно определять связь между различными лингвистическими и металингвистическими феноменами.')
``` 
